A 3d curvilinear skeletonization algorithm with application to path tracing

### 2012-09-11 application_to_path_tracing.tex
Renommage fichier tex application

Let $O$ be the origin of the camera in the scene. For each pixel $P$ of the final image we search for the nearest intersection point $x$ of the ray $(O, \overrightarrow{OP} = -\Theta)$ with the scene. At this point, to obtain the luminosity, we must solve the rendering equation \cite{Ka86}. The rendering equation is a recursive integral equation. The integrand of the equation contains the radiance function that we must compute:

@@

$L$ is the radiance (a radiometric quantity that represents luminosity) from point $x$ toward direction $\Theta$. $L_e$ is the emitted radiance that is not zero only on light sources. $r(x, \Phi)$ is the nearest visible point from $x$ in the direction $\Phi$. The $f_r$ function is called the \myem{bidirectional reflectance function} (BRDF) and express the exchange of radiance between the incoming direction $\Phi$ and the output direction $\Theta$. Intuitively it represents the material properties of the surface at point $x$. A mirror doesn't reflect light the same manners a wall do, the BRDF encodes that.

This equation expresses an intuitive idea: the radiance $L$ produced by the point $x$ in the direction $\Theta$ is the sum of the emitted radiance and the reflected radiance.  The reflected radiance is computed by integrating the incoming radiance scaled by a factor $f_r(x, \Theta \leftrightarrow \Phi) cos(N_x, \Phi)$ over all the possible directions of the hemisphere around $x$. $N_x$ is the surface normal at point $x$.

In practice it is impossible to compute anatically the integral, it has to be estimated.
The most commonly used method is the Monte-Carlo integration that provides the following estimator for the integral:
\begin{equation*}
\langle L_r(x \rightarrow \Theta) \rangle = \frac{f_r(x, \Theta \leftrightarrow \Phi) L(r(x, \Phi) \rightarrow -\Phi) cos(N_x, \Phi)}{p(\Phi)}
\end{equation*}

$p$ is a probability density function (pdf) that is used to sample $\Phi$, a direction of the hemisphere $\Omega_x$. This estimator is unbiased, it means that the expected value $E[\langle L_r(x \rightarrow \Theta) \rangle]$ is equal to $L_r(x \rightarrow \Theta)$. The variance of the estimator express its quality.
The variance of the estimator depends on the choosen pdf $p$. Actually the best is to choose a pdf that matches the shape of the function to integrate (ie. give high density to samples that have high values for the function and low density to samples that have low values).

@@

The implantation of $L$ is as follow: first we sample a random direction $\Phi$ based on the pdf $p$. Then we apply the estimator that call $L$ recursively to compute $L_r$. Then we return the sum of the emitted radiance $L_e$ and the result for $L_r$. The recursion stops when a maximal number of bounces have been reach.

This algorithm is called for each pixel of the final image to compute. To improve the quality of the final image we can throw multiple rays accross each pixel and compute the mean of the results.

In the radiance algorithm the $\Phi$ direction is sampled on $\Omega_x$ based on the pdf $p$. A bad pdf picks directions that don't reach the light before the end of the recursion and results in a lot of noise in the final image. We will discuss the choice of $p$ in the next section.

@@

The key idea behind our method is that light travels in the void of the scene. A curvilinear skeleton of that void gives us enough informations to compute some importance directions that tells us where the light comes from. Given that directions we can construct a efficient pdf $p_{skel}$ and guide our rays by sampling the hemispheres with $p_{skel}$

Here is the integrand of the integral estimated with Monte-Carlo:

\begin{equation*}
f_r(x, \Phi, \Theta) L(x \leftarrow \Phi) cos(N_x, \Phi)
\end{equation*}

@@

$L$ is rarelly used because we cannot evaluate it in constant time. Remember that it represents the distribution of light in the scene. It takes high values when $\Phi$ carries a lot of energy. Out method gives a way to sample according to $L$.

@@

The skeleton of the void is computed with our algorithm and is converted to a graph (the nodes are 3D positions and the edges represents the topology of the scene), given as an input of our path tracing. We then compute a set of points called \myem{importance points}. These points will be used to sample $\Omega_x$ in the path tracing algorithm. For each node of the skeleton $n$, one importance point $imp_n$ is computed.

@@

Let $n$ be a node of the skeleton and $V_n$ the set of visible nodes from $n$ along the shortest path toward $n_L$. The importance point $imp_n$ associated to $n$ is computed by taking the barycenter of $V_n$.

@@

Given a point $x$ and a direction $\Theta$ we want to compute $L(x \rightarrow \Theta)$ and then sample the hemisphere $\Omega_x$. We search for the nearest skeleton node $n$ and its importance point $imp_n$. We sample the hemisphere with a power-cosine pdf center on $\overrightarrow{ximp_n}$:

@@

with $\alpha$ the angle between $\overrightarrow{ximp_n}$ and $\Phi$ and $s$ a parameter called skeleton strength. The more $s$ is high, the more we sample close $\overrightarrow{ximp_n}$.

@@

We present the results of our algorithm applied to four scenes. the first row shows our result and the second row shows a standard path tracing with the same input parameters. On each picture a MSE (mean square error) value with a reference image is shown.

\begin{center}
\hspace*{-40px}
\includegraphics[scale=0.2]{images/four_scenes.png}
\end{center}

We observe a net decreasing of the noise for our algorithm, especially in dark areas. The MSE values are lower for the top pictures, meaning that we converge faster to the reference image than standard path tracing does.

We presented an application of our skeletonization algorithm for global illumination. Our strategy works well, it reduces noise by taking into account the distribution of energy in the scene. One of the problem is that we take into account only one importance direction for sampling, we totally ignore the BRDF and the cosine factor. A solution called Multiple Importance Sampling allow us to combine different strategies into one. We will not describe this strategy because it's not directly related to skeletonization, but it has allowed us to improve our results and remove some artifacts of our method.

>>>>

Let $O$ be the origin of the camera in the scene. For each pixel $P$ of the final image we search for the nearest intersection point $x$ of the ray $(O, \overrightarrow{OP} = -\Theta)$ with the scene. At this point, to obtain the luminosity, we must solve the rendering equation \cite{Ka86} which is a recursive integral equation. The integrand of the equation contains the radiance function that we must compute:

@@

$L$ is the radiance (a radiometric quantity that represents luminosity) from point $x$ toward direction $\Theta$. $L_e$ is the emitted radiance that is not zero only on light sources. $r(x, \Phi)$ is the nearest visible point from $x$ in the direction $\Phi$ of the hemisphere $\Omega_x$. $N_x$ is the surface normal at point $x$. The $f_r$ function is called the \myem{bidirectional reflectance function} (BRDF) and express the exchange of radiance between the incoming direction $\Phi$ and the output direction $\Theta$. Intuitively it represents the material properties of the surface at point $x$. A mirror does not reflect light the same manners a wall do and the BRDF describes that kind of behavior.

This equation expresses an intuitive idea: the radiance $L$ produced by the point $x$ in the direction $\Theta$ is the sum of the emitted radiance and the reflected radiance.  The reflected radiance is computed by integrating the incoming radiance scaled by a factor $f_r(x, \Theta \leftrightarrow \Phi) cos(N_x, \Phi)$ over all the possible directions of the hemisphere around $x$.

The most commonly used method to compute the integral is the Monte-Carlo integration that provides the following estimator:
\begin{equation*}
\langle L_r(x \rightarrow \Theta) \rangle = \frac{f_r(x, \Theta \leftrightarrow \Phi) L(r(x, \Phi) \rightarrow -\Phi) cos(N_x, \Phi)}{p(\Phi)}
\end{equation*}

where $p$ is a probability density function (pdf) that is used to sample $\Phi$. This estimator is unbiased, meaning that the expected value $E[\langle L_r(x \rightarrow \Theta) \rangle]$ is equal to $L_r(x \rightarrow \Theta)$. The variance of the estimator expresses its quality and depends on the choosen pdf $p$. Actually the best is to choose a pdf that matches the shape of the function to integrate (ie. give high density to samples that have high values for the function and low density to samples that have low values).

@@

The algorithm to compute $L$ is as follow: first we sample a random direction $\Phi$ based on the pdf $p$. Then we apply the estimator that call $L$ recursively to compute $L_r$. Finally we return the sum of the emitted radiance $L_e$ and the result for $L_r$. The recursion stops when a maximal number of bounces has been reach.

This computation is done multiple time for each pixel (at different positions over the pixel). We average the results and get an estimation of the mean radiance passing through the pixel in direction of the camera's origin.

In the radiance computation the $\Phi$ direction is sampled on $\Omega_x$ based on the pdf $p$. A bad pdf picks directions that don't reach the light before the end of the recursion and results in a lot of noise in the final image. We will discuss the choice of $p$ in the next section.

@@

The key idea behind our method is that light travels in the void of the scene. As stated in the introduction, a curvilinear skeleton of the void gives us enough informations to compute some importance directions where light comes from. Given these directions we can construct a efficient pdf $p_{skel}$ and guide our rays by sampling the hemispheres with $p_{skel}$

The integrand of $L_r$ is:

\begin{equation*}
f_r(x, \Theta \leftrightarrow \Phi) L(x \leftarrow \Phi) cos(N_x, \Phi)
\end{equation*}

@@

$L$ is rarelly used because its evaluation cannot be done in constant time. Remember that it represents the distribution of light in the scene. It takes high values when $\Phi$ carries a lot of energy. Our method gives a way to sample according to $L$.

@@

The skeleton of the void is computed with our algorithm \ref{sec::skeletonization} and is converted to a graph (the nodes are 3D positions and the edges represents the topology of the scene), given as an input of our path tracing. We then compute a set of points called \myem{importance points}. These points will be used to sample $\Omega_x$ in the path tracing algorithm. For each node of the skeleton $n$, one importance point $imp_n$ is computed.

@@

Let $n$ be a node of the skeleton and $V_n$ the set of visible nodes from $n$ along the shortest path toward $n_L$. The importance point $imp_n$ associated to $n$ is the barycenter of $V_n$.

@@

Given a point $x$ and a direction $\Theta$ we want to compute $L(x \rightarrow \Theta)$ and then sample the hemisphere $\Omega_x$. We search for the nearest skeleton node $n$ and its importance point $imp_n$. We sample the hemisphere with a power-cosine pdf centered on $\overrightarrow{ximp_n}$:

@@

with $\alpha$ the angle between $\overrightarrow{ximp_n}$ and $\Phi$, $s$ being a parameter called skeleton strength. The higher $s$ is, the more we sample close $\overrightarrow{ximp_n}$.

@@

We present some results of our algorithm applied to four scenes. the first row shows our result and the second row shows a standard path tracing with the same input parameters. On each picture a MSE (mean square error) value with a reference image is shown.

\begin{center}
\includegraphics[scale=0.17]{images/four_scenes.png}
\end{center}

We observe a decrease of the noise for our algorithm, especially in dark areas. The MSE values are lower for the top pictures, meaning that we converge faster to the reference image than standard path tracing does.

We presented an application of our skeletonization algorithm for global illumination. Our strategy works well, it reduces noise by taking into account the distribution of energy in the scene. One of the problem is that we take into account only one importance direction for sampling, totally ignoring the BRDF and the cosine factor. A solution called Multiple Importance Sampling will allow us to combine different strategies into one.
